---
title: "01 - Adding a plot and control"
author: "Ted Laderas"
---

```{r setup, include=FALSE}
library(shiny)
library(ggplot2)
library(fivethirtyeight)
library(vembedr)
data(biopics)
```

# Slides

https://laderast.github.io/gradual_shiny/slides/#10

# Video

```{r}
vembedr::embed_url("https://youtu.be/_UuuL-cFqjs?t=170")
```


## Starting with a minimal shiny app

The code below is the bare minimum you need to start a Shiny app. 

To run it, you can click the "play" button. 


```{r}
library(shiny)

ui <- fluidPage()  #ui element, handles controls and displaying plots

server <- function(input, output) {} #server element, handles calculations

shinyApp(ui = ui, server = server)
```

We'll need to populate both of the `ui` and `server` objects. 


## `input` and `output` are fixed, but not what goes in them.

The `input` and `output` objects 



## Adding a `ggplot` to our app

The first thing we're going to do is add a `ggplot()` command to our app. 

To do this, we'll add the following plot to our shiny app. 

Why do we use `aes_string()` instead of `aes()`? With `aes_string()`, we can specify variables as `character`, which will let us programmatically change these variables on the fly.

```{r}
    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
```
This part goes into the server portion of our app. So that `ui` can access it, we'll assign it to `output$paired_plot`, so that our `ui` can "see" it. Remember, `server` controls the `output` object.

```
server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}
```

We're not done yet! To display it, we also need to add a `plotOutput()` to our `ui`:

```
ui <- fluidPage(
  plotOutput("paired_plot")
)
```

## Question

Where did the "paired_plot" come from in the `ui` element? 


## Shiny App with a Plot

Here's the new version of the app now. It will just display our plot, but we're now going to modify it.

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)

ui <- fluidPage(
  plotOutput("paired_plot")
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```





## 1.3 Adding a select box to our app

Let's add a control called `selectInput()` to our app. It goes after our `plotOutput` in our `ui` element:

```
ui <- fluidPage(
  plotOutput("paired_plot"),  # <--- note the comma here!
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

```

Note that it's not yet wired to control `paired_plot` yet. 

We've just added the control to our `ui`.

### Question:

Where does `categoricalVars` come from? Where is it defined in the app?

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  plotOutput("paired_plot"),
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```


## Connecting our `selectInput()` to `server` using the `input` object.

Now we need to connect our `selectInput` to our `ggplot()`. Specifically, we will use the value of our `selectInput()` to control the `color` aesthetic.


## Exercise

Wire the `selectInput()` control to the `color` aesthetic of the `ggplot()` statement by adding `input$color_select` (you'll add it where the blank is).

Verify that you have wired it correctly by running the app code and seeing whether the select box controls the color of the plots.

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars),
  plotOutput("paired_plot")
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= _______________) +   #fill in this blank
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

Where did `input$color_select` variable come from?



## Mistakes I make quite often

The `ui` element expects each of the `_Input()`s and `_Output()`s separated by commas. So, if your code isn't running and ui is giving you an error, check for a missing comma!

## Final App

This is also available in the `app.R` file in the `01_app_basics` folder. 

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  plotOutput("paired_plot"),
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= input$color_select) +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

## Let's go on!