

## Starting with a minimal shiny app

The code below is the bare minimum you need to start a Shiny app. You can run it, but it won't do anything.

```{r}
library(shiny)

ui <- fluidPage()  #ui element, handles controls and displaying plots

server <- function(input, output) {} #server element, handles calculations

shinyApp(ui = ui, server = server)
```

## `input` and `output` are fixed, but not what goes in them.





## Adding a `ggplot` to our app

The first thing we're going to do is add a `ggplot()` command to our app. 

To do this, we'll add the following plot to our shiny app. 

Why do we use `aes_string()` instead of `aes()`? With `aes_string()`, we can specify variables as `character`, which will let us programmatically change these variables on the fly.

```{r}
    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
```
This part goes into the server portion of our app. So that `ui` can access it, we'll assign it to `output$paired_plot`, so that our `ui` can "see" it. Remember, `server` controls the `output` object.

```
server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}
```

We're not done yet! To display it, we also need to add a `plotOutput()` to our `ui`:

```
ui <- fluidPage(
  plotOutput("paired_plot")
)
```

## Question

Where did the "paired_plot" come from in the `ui` element? 


## Shiny App with a Plot

Here's the new version of the app now. It will just display our plot, but we're now going to modify it.

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)

ui <- fluidPage(
  plotOutput("paired_plot")
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

## Adding a select box to our app

Let's add a control called `selectInput()` to our app. It goes after our `plotOutput` in our `ui` element:

```
ui <- fluidPage(
  plotOutput("paired_plot"),  # <--- note the comma here!
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

```

Note that it's not yet wired to control `paired_plot` yet. 

We've just added the control to our `ui`.

### Question:

Where does `categoricalVars` come from? Where is it defined in the app?

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  plotOutput("paired_plot"),
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

## Connecting our `selectInput()` to `server` using the `input` object.

Now we need to connect our `selectInput` to our `ggplot()`. Specifically, we will use the value of our `selectInput()` to control the `color` aesthetic


## Exercise

Wire the `selectInput()` control to the `color` aesthetic of the `ggplot()` statement by adding `input$color_select` (you'll add it where the blank is).

Verify that you have wired it correctly by running the app code and seeing whether the select box controls the color of the plots.

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars),
  plotOutput("paired_plot")
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= _______________) +   #fill in this blank
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

Where did `input$color_select` variable come from?


## Mistakes I make quite often

The `ui` element expects each of the `_Input()`s and `_Output()`s separated by commas. So, if your code isn't running and ui is giving you an error, check for a missing comma!

## Final App

This is also available in the `app.R` file in the `01_app_basics` folder. 

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  plotOutput("paired_plot"),
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= input$color_select) +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

# Making a `reactive()` dataset for our app.

So far, we've just changed aesthetics for the plot in our app.

What if we wanted to change the underlying data using `filter()`? 

Can we dynamically control that process with a control on the `shiny()` app?

Yes, we can, with a new idea called a `reactive`.


## What is a Reactive?

Much like our plot was listening for changes to our `selectInput()` control, we can make a dataset that listens for changes to a `slider`. 



## Adding a slider 

Let's go back to the simple form of our app. We'll take out the `selectInput()` for right now, so we can focus on our new slider and what it can do.

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)

ui <- fluidPage(
  plotOutput("paired_plot"),
  sliderInput(inputId = "year_slider",
              label = "Select Year to Filter",
              min = 1915,
              max = 2014,
              value = 1915)
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

## Making our Dataset react to our slider.

We're going to take the following `dplyr` code and make it react to our slider:


```{r}
biopics %>%
  filter(year > 1915)
```

To do this, we need to first put it into what's called a `reactive()`.


```
biopics_filtered <- reactive({
   biopics %>%
      filter(year > 1915)
})

```
## Final App

Here is the final app (also available in `02_reactives/app.R`).

```{r}
library(shiny)
library(tidyverse)
library(fivethirtyeight)

##load the biopics data
data(biopics)
biopics <- biopics %>% filter(!is.na(box_office))

# Define UI for application that plots 
ui <- fluidPage(
   
      plotOutput("scatter_plot"),
      sliderInput("year_filter", 
                  "Select Lowest Year", 
                  min = 1915,
                  max=2014, 
                  value = 1915)
         
)

##Server is where all of the computations happen
server <- function(input, output) {
  
  biopics_filtered <- reactive({
    biopics %>%
        filter(year_release > input$year_filter)
     })
   
   output$scatter_plot <- renderPlot({
      
       ggplot(biopics_filtered()) +
         aes_string(y="box_office", 
         x="year_release") + 
       geom_point()
     
   })
}

# Run the application 
shinyApp(ui = ui, server = server)
```