```{r}
library(shiny)
library(vembedr)
library(tidyverse)
library(fivethirtyeight)
data("biopics")
```

# Slides about Reactives

https://laderast.github.io/gradual_shiny/slides/#27


# Video

```{r}
vembedr::embed_url("https://youtu.be/_UuuL-cFqjs?t=3080")
```


# Making a `reactive()` dataset for our app.

So far, we've just changed aesthetics for the plot in our app.

What if we wanted to change the underlying data using `filter()`? 

Can we dynamically control that process with a control on the `shiny()` app?

Yes, we can, with a new idea called a `reactive`.



## What is a Reactive?

Much like our plot was listening for changes to our `selectInput()` control, we can make a dataset that listens for changes to a `slider`. 




## Adding a slider 

Let's go back to the simple form of our app. We'll take out the `selectInput()` for right now, so we can focus on our new slider and what it can do.

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)

ui <- fluidPage(
  plotOutput("paired_plot"),
  sliderInput(inputId = "year_slider",
              label = "Select Year to Filter",
              min = 1915,
              max = 2014,
              value = 1915)
)

server <- function(input, output) {

  output$paired_plot <- renderPlot({

    ggplot(biopics) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

## Making our Dataset react to our slider.

We're going to take the following `dplyr` code and make it react to our slider:


```{r}
biopics %>%
  filter(year < 2014)
```

To do this, we need to first put it into what's called a `reactive()`.


```
biopics_filtered <- reactive({
   biopics %>%
      filter(year < 2014)
})

```

## Exercise

Wire the sliderInput to the reactive by using the correct id to `input` (it's the blank below.)

When you are done with that, use `biopics_filtered()` as a dataset in the ggplot statement:

```{r}
library(shiny)
library(fivethirtyeight)
data(biopics)

ui <- fluidPage(
  plotOutput("paired_plot"),
  sliderInput(inputId = "year_slider",
              label = "Select Year to Filter",
              min = 1915,
              max = 2014,
              value = 1915)
)

server <- function(input, output) {
  
  biopics_filtered <- reactive({
    
    biopics %>%
      filter(year < input$_______)
    
  })
  

  output$paired_plot <- renderPlot({

    ggplot(___________) + 
      aes_string(x="year_release", 
                 y="box_office", 
                 color= "type_of_subject") +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```



## Final App

Here is the final app (also available in `02_reactives/app.R`).

```{r}
library(shiny)
library(tidyverse)
library(fivethirtyeight)

##load the biopics data
data(biopics)
biopics <- biopics %>% filter(!is.na(box_office))

# Define UI for application that plots 
ui <- fluidPage(
   
      plotOutput("scatter_plot"),
      sliderInput("year_filter", 
                  "Select Highest Year",
                  min = 1915,
                  max=2014, 
                  value = 2014)
         
)

##Server is where all of the computations happen
server <- function(input, output) {
  
  biopics_filtered <- reactive({
    biopics %>%
        filter(year_release < input$year_filter)
     })
   
   output$scatter_plot <- renderPlot({
      
       ggplot(biopics_filtered()) +
         aes_string(y="box_office", 
         x="year_release",
         color="type_of_subject") + 
       geom_point()
     
   })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

