---
title: '03: Plotly'
author: "Ted Laderas"
date: "1/27/2021"
output: html_document
---

```{r setup, include=FALSE}
library(shiny)
library(fivethirtyeight)
library(ggplot2)
library(plotly)
library(vembedr)
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

```

# Slides

https://laderast.github.io/gradual_shiny/slides/#42

# Video

```{r}
vembedr::embed_url("https://youtu.be/_UuuL-cFqjs?t=4881")
```

# Converting a ggplot to plotly using `ggplotly()`


```{r}

my_plot <- ggplot(biopics) +
      aes_string(y = "box_office", 
                 x="year_release",
                 color="type_of_subject") +
      geom_point() +
      theme(legend.position="none")

my_plot
```

You can convert `my_plot` to a plotly object using the ggplotly function:

```{r}
ggplotly(my_plot)
```

You can also control which aesthetics are shown using the `tooltip` argument:

```{r}
ggplotly(my_plot, tooltip = c("box_office", "year_release"))

```

Now to add this to an app, we can add our plotly object to an app using `plotlyOutput()` for `ui`, and `renderPlotly({})` for `server`:

```{r}
ui <- fluidPage(
  plotlyOutput("movie_plot")
)

server <- function(input, output) {
  
  output$movie_plot <- renderPlotly({
    
    my_plot <- ggplot(biopics) +
      aes_string(x = "box_office", 
                 y="year_release",
                 color="type_of_subject") +
      geom_point() +
      theme(legend.position="none")
    
    ##notice we pass our ggplot into ggplotly, which makes it more interactive
    ggplotly(my_plot)
  })

}

shinyApp(ui, server)
```

## Exercise

1. Add more aesthetics to your ggplot `aes_string()` statement, such as movie_title="title". Here's a list of the variables in `biopics` for your reference.

```{r}
colnames(biopics)
```

2. Then select a subset of these aesthetics to display using the `tooltip` argument to `ggplotly()`.


```{r}
ui <- fluidPage(
  plotlyOutput("movie_plot")
)

server <- function(input, output) {
  
  output$movie_plot <- renderPlotly({
    
    my_plot <- ggplot(biopics) +
      aes_string(x = "box_office", 
                 y="year_release",
                 color="type_of_subject",
                 ----------------------,
                 ----------------------,) +
      geom_point() +
      theme(legend.position="none")
    
    ##notice we pass our ggplot into ggplotly, which makes it more interactive
    ggplotly(my_plot, tooltip = c(---------------------))
  })

}

shinyApp(ui, server)
```


# Final App

Here's the final app using our new `plotly` plot.


```{r}
data(biopics)
categoricalVars <- c("country", "type_of_subject", "subject_race", "subject_sex")

ui <- fluidPage(
  plotlyOutput("movie_plot"),
  selectInput(inputId = "color_select", 
              label = "Select Categorical Variable", 
              choices = categoricalVars)
)

server <- function(input, output) {
  
  output$movie_plot <- renderPlotly({
    
    my_plot <- ggplot(biopics) +
      aes_string(x = "box_office", 
                 y="year_release",
                 color=input$color_select,
                 title="title",
                 director="director",
                 box_office="box_office", 
                 subject="subject") +
      geom_point() +
      theme(legend.position="none")
    
    ##notice we pass our ggplot into ggplotly, which makes it more interactive
    ggplotly(my_plot)
  })

}

shinyApp(ui, server)
```
