---
title: "a gRadual intRoduction to Shiny"
description: |
  Introductory Workshop to Basic Shiny Concepts
author:
  - first_name: "Ted"
    last_name: "Laderas"
    url: https://laderast.github.io
    affiliation: Division of Bioinformatics and Computational Biomedicine, Medical Informatics and Clinical Epidemiology; Knight Cancer Institute, Oregon Health & Science University
  - first_name: "Jessica"
    last_name: "Minnier"
    url: https://jessicaminnier.com
    affiliation: PSU/OHSU School of Public Health; Knight Cancer Institute, Oregon Health & Science University
  - first_name: "Pierrette"
    last_name: "Lo"
    affiliation: Knight Cancer Institute, Oregon Health & Science University
    url: https://lopierra.rbind.io/
  - first_name: "Dar'ya"
    last_name: "Pozhidayeva"
    affiliation: Oregon Clinical Translational Research Institute; Medical Informatics and Clinical Epidemiology, Oregon Health & Science University.
date: January 28, 2018
site: distill::distill_website
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html
library(metathis)
library(vembedr)
```

```{r}
meta() %>%
  meta_description(
    "This workshop is a gentle introduction to Shiny and related concepts"
  ) %>% 
  meta_name("github-repo" = "laderast/gradual_shiny") %>% 
  meta_viewport() %>% 
  meta_social(
    title = "a gRadual intRoduction to Shiny",
    url = "https://laderast.github.io/gradual_shiny",
    og_type = "Workshop",
    og_author = c("Ted Laderas", "Jessica Minnier"),
    twitter_card_type = "summary",
    twitter_creator = "@tladeras"
  )
```


# Description

This was a two hour workshop given for [OHSU BioData Club](https://biodata-club.github.io) on 2021-01-28 and Cascadia-R 2021-06-04 introducing the basic concepts of Shiny. It runs about two hours long.

There are videos for each part, along with slides. We use an RStudio.cloud project to try out the basic Shiny concepts.

# Audience

Intermediate R users and students. Students should be familiar with `{ggplot2}`, `{dplyr}`, and how functions work.

# Learning Objectives

- **Learn** the basic architecture of Shiny Apps
- **Learn** how `server` and `ui` communicate with the `input` and `output` objects
- **Add** `ggplot2` code to our app and display it with `plotOutput` and `renderOutput`
- **Add** a control to control the aesthetics of our `ggplot`
- **Make** a reactive dataset using reactive expressions and control it with a slider
- **Use** `{plotly}` to add tooltips to our data
- **Learn** about the Shiny Widget Gallery and how to use it to add controls to your app


# RStudio Cloud Project

To use this project, you will need to sign up for an https://rstudio.cloud free account. Make sure to [save a permanent copy to your own workspace](https://rstudio.cloud/learn/guide#your-workspace)!

https://rstudio.cloud/project/2125039  

# Part 1: Basic Architecture of a Shiny App

## Slides

Open slides in a separate window: <a href="https://laderast.github.io/gradual_shiny/slides/#10" target="_blank">https://laderast.github.io/gradual_shiny/slides/#10</a>


```{r}
knitr::include_url("https://laderast.github.io/gradual_shiny/slides/#10")
```

## Video

https://youtu.be/6XBb5J6U05Y?t=2956

```{r}
vembedr::embed_youtube("6XBb5J6U05Y") 
```

## Project

Open up `gradual_shiny/01_app_basics/01_app_basics.Rmd` in RStudio Project


# Part 2: Reactives

## Slides

Open slides in a separate window: <a href="https://laderast.github.io/gradual_shiny/slides/#27" target="_blank">https://laderast.github.io/gradual_shiny/slides/#27</a>


```{r}
knitr::include_url("https://laderast.github.io/gradual_shiny/slides/#27")
```

## Video

```{r}
vembedr::embed_youtube("6XBb5J6U05Y") %>%
  vembedr::use_start_time("49m16")
```

## Project

Open up `gradual_shiny/02_reactives/02_reactives.Rmd` in RStudio Project

# Part 3: Tooltips with `{plotly}`; Wrapping up

## Slides

Open slides in a separate window: <a href="https://laderast.github.io/gradual_shiny/slides/#37" target="_blank">https://laderast.github.io/gradual_shiny/slides/#36</a>


```{r}
knitr::include_url("https://laderast.github.io/gradual_shiny/slides/#37")
```

## Video

```{r}
vembedr::embed_youtube("_UuuL-cFqjs") %>%
  vembedr::use_start_time("72m48")
```

## Project

Open up `gradual_shiny/03_tooltips_plotly/03_tooltips_plotly.Rmd` in RStudio Project.

# Survey

If you've taken the workshop, please take our survey and give us feedback: https://ohsu.ca1.qualtrics.com/jfe/form/SV_afaOiEAnNpWZIoK

If you want to use the workshop materials, please drop us a line at [tedladeras] [at] [gmail]. We'd love to hear from you.
