---
title: "a gRadual intRoduction to Shiny"
description: |
  Introductory Workshop to Basic Shiny Concepts
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html
library(metathis)
library(vembedr)
```

```{r}
meta() %>%
  meta_description(
    "This workshop is a gentle introduction to Shiny and related concepts"
  ) %>% 
  meta_name("github-repo" = "laderast/gradual_shiny") %>% 
  meta_viewport() %>% 
  meta_social(
    title = "a gRadual intRoduction to Shiny",
    url = "https://laderast.github.io/gradual_shiny",
    og_type = "Workshop",
    og_author = c("Ted Laderas", "Jessica Minnier"),
    twitter_card_type = "summary",
    twitter_creator = "@tladeras"
  )
```


# Description

This was a two hour workshop given for BioData Club on 2021-01-28 introducing the basic concepts of Shiny. It runs about two hours long.

There are videos for each part, along with slides. We use an RStudio.cloud project to try the basic Shiny concepts.

# Audience

Intermediate R users and students. Students should be familiar with `{ggplot2}`, `{dplyr}`, and how functions work.

# Learning Objectives

- **Learn** the basic architecture of Shiny Apps
- **Learn** how `server` and `ui` communicate with the `input` and `output` objects
- **Add** `ggplot2` code to our app and display it with `plotOutput` and `renderOutput`
- **Add** a control to control the aesthetics of our `ggplot`
- **Make** a reactive dataset using reactive expressions and control it with a slider
- **Use** `{plotly}` to add tooltips to our data
- **Learn** about the Shiny Widget Gallery and how to use it to add controls to your app


# RStudio Cloud Project

Make sure to save a permanent copy to your own workspace!

https://rstudio.cloud/project/2125039  

# Part 1: Basic Architecture of a Shiny App

## Slides

Open slides in a separate window: <a href="https://laderast.github.io/gradual_shiny/slides/#10" target="_blank">https://laderast.github.io/gradual_shiny/slides/#10</a>


```{r}
knitr::include_url("https://laderast.github.io/gradual_shiny/slides/#10")
```

## Video

```{r}
vembedr::embed_youtube("_UuuL-cFqjs") %>%
  use_start_time("2m48")
```

## Project

Open up `gradual_shiny/01_app_basics/01_app_basics.Rmd` in RStudio Project


# Part 2: Reactives

## Slides

Open slides in a separate window: <a href="https://laderast.github.io/gradual_shiny/slides/#27" target="_blank">https://laderast.github.io/gradual_shiny/slides/#27</a>


```{r}
knitr::include_url("https://laderast.github.io/gradual_shiny/slides/#27")
```

## Video

```{r}
vembedr::embed_youtube("_UuuL-cFqjs") %>%
  vembedr::use_start_time("51m20")
```

## Project

Open up `gradual_shiny/02_reactives/02_reactives.Rmd` in RStudio Project

# Part 3: Tooltips with `{plotly}`; Wrapping up

## Slides

Open slides in a separate window: <a href="https://laderast.github.io/gradual_shiny/slides/#36" target="_blank">https://laderast.github.io/gradual_shiny/slides/#36</a>


```{r}
knitr::include_url("https://laderast.github.io/gradual_shiny/slides/#36")
```

## Video

```{r}
vembedr::embed_youtube("_UuuL-cFqjs") %>%
  vembedr::use_start_time("81m21")
```

## Project

Open up `gradual_shiny/03_tooltips_plotly/03_tooltips_plotly.Rmd` in RStudio Project.

# Survey

If you've taken the workshop, please take our survey and give us feedback: https://ohsu.ca1.qualtrics.com/jfe/form/SV_afaOiEAnNpWZIoK

If you want to use the workshop materials, please drop us a line at [tedladeras] [at] [gmail]. We'd love to hear from you.
